% @author Huyen Tran DONE

\noindent This book provides a comprehensive exploration of the Maximum Matching problem in graph theory, beginning with foundational principles and progressing to advanced algorithmic techniques and applications. It builds on fundamental concepts such as graph structure, adjacency, and subgraphs to create a solid theoretical basis for understanding matching problems and their inherent complexity.

\noindent From this foundation, the book delves into a wide range of algorithms tailored to specific graph types, including bipartite, d-partite, and general graphs. It examines the efficiency and practicality of methods such as Hopcroft-Karp, Edmonds' Blossom algorithm, and heuristic approaches like Ant Colony Optimization, offering both exact solutions and scalable approximations for challenging instances.

\noindent Beyond algorithmic details, the book explores the connections between maximum matching and related optimization problems, such as perfect matching, vertex cover, and maximum flow. These relationships are presented through reductions, dual formulations, and polytope characterizations, providing a broader perspective on the role of matching in graph theory and optimization.

\noindent Finally, the book highlights practical implementations and real-world applications, demonstrating how maximum matching algorithms are employed in network design, scheduling, and other domains. By combining theoretical rigor with practical relevance, this book equips readers with the tools to analyze, implement, and innovate within the field of graph optimization. Whether for academic study or applied problem-solving, it serves as a comprehensive guide to the Maximum Matching problem and its broader implications in computational science.